---
title: "Homework 6"
author: "Pooja Mukund"
date: "11/26/2021"
output: github_document
---

```{r}
#Load relevant libraries

library(tidyverse)
library(viridis)
```

```{r}
#Setup for visualizations
knitr::opts_chunk$set(
  fig.width = 10,
  fig.asp = .6,
  out.width = "90%"
)
```

## Load Data
```{r}
birth<- read_csv("data/birthweight.csv")
```

## Data Cleaning   

First we will check if there are any NAs in columns 
```{r}
# Check for NAs 
birth%>%
summarise_all(funs(sum(is.na(.)))) # No NAs in columns 
```
From the data description in homework assignment we know that `babysex`, `frace`, `malform`, `mrace` are all categorical variables that are currently coded numerically. We will first change this to categorical variables.  

```{r}
birth%>%
  summarise_all(class)
#All coded numerically 

birth%>%
  distinct(frace)
#Only 5 categories - 1, 2, 3, 4, 8 even though description has 9 there are no 9s that exist in this current dataset 

birth%>%
  distinct(mrace)
#Only 1, 2, 3, 4 for mrace

birth%>%
  distinct(malform)

birth_cln<-birth%>%
  mutate( babysex = replace(babysex, babysex == 1, "Male"),
          babysex = replace(babysex, babysex == 2, "Female"), 
          frace = replace(frace, frace ==1, "White"),
          frace = replace(frace, frace ==2, "Black"),
          frace = replace(frace, frace ==3, "Asian"),
          frace = replace(frace, frace ==4, "Puerto Rican"),
          frace = replace(frace, frace ==8, "Other"),
          malform = replace(malform, malform ==0, "absent"), 
          malform = replace(malform, malform ==1, "present"),
          mrace = replace(mrace, mrace ==1, "White"),
          mrace = replace(mrace, mrace ==2, "Black"),
          mrace = replace(mrace, mrace ==3, "Asian"),
          mrace = replace(mrace, mrace ==4, "Puerto Rican"), 
          babysex = factor(babysex, levels=c("Male", "Female")), 
          frace = factor(frace, levels=c("White", "Black", "Asian", "Puerto Rican", "Other")),
          mrace = factor(mrace, levels=c("White", "Black", "Asian", "Puerto Rican")))

birth_cln%>%
  head(5)%>%
  knitr::kable()
```

Regression model for birthweight

From this [publication](https://pubmed.ncbi.nlm.nih.gov/7570074/), I hypothesize that maternal race, infant sex, maternal smoking status, and education would be important factors for birthweight. Since we do not have education, I will use family monthly income (`fincome`) as a proxy for education. Who defines smoking status as "someone who smokes a tobacco product either daily or occasionally". I will use `smoken` and create a new variable `smoke_status` defined by smoker as greater than 0 cigrattes per day and non-smoker as 0 cigarettes per day.  


```{r}
#Create smoker variable 
birth_cln%>%
  distinct(smoken)

birth_cln<-birth_cln%>%
  mutate(smoke_status = case_when(smoken >0 ~ "smoker",
                                  smoken ==0 ~"non-smoker"))
```


Model Fitting Process - We will start with `smoke_status` and `babysex` and evaluate based on Adjusted $R^2$ 
```{r}
fit = lm(bwt ~ smoke_status + babysex, data = birth_cln)
summary(fit)
```

Let's see if adding maternal race helps this model 
```{r}
fit_2 = lm(bwt ~ smoke_status + babysex + mrace, data = birth_cln)
summary(fit_2)
```
This slightly increased our Adjusted $R^2$ from 0.01957 to 0.1149. Let's see if adding family monthly income helps (`fincome`). 

```{r}
fit_3 = lm(bwt ~ smoke_status + babysex + mrace+fincome, data = birth_cln)
summary(fit_3)
```

This slightly increased our Adjusted $R^2$ from 0.1149 to 0.1157. Still not the best model, this may suggest a linear model is not the best fit for birthweight. However, we will continue using `fit_3`. 

Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r}

```


## Problem 2 
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

